---
title: "cell-counts"
author: "ML Larsen"
date: "Saturday, January 17, 2015"
output: pdf_document
---

## Background
Counting the cellular density of cultures can be quite involved. This document takes you through a step-by-step process from sample and filter prep to calculating the density in a culture. R scripts can be used for single sampling or across a time course.

```{r}
#remove variable assignments from RStudio environment
rm(list = ls())

#set your working directory
setwd("C:/Users/Megan/GitHub//cell-counts")
```

## suggestions for R code
* single use excel template with easy to use calculation setup
* R calibration file for filter settings

## Scope calibrations
Inside diameter of filter tower = 1.64 cm; Area of filter = 2.1 cm^2

Dimensions of field calculated by capturing image of slide micrometer at each magnification

x20 = 0.0013 cm2 
x40 = 0.0003 cm2 
x1000 = 0.0001 cm2

```{r}
## Number of fields for each objective; DO NOT CHANGE THESE VALUES

x20 <- 1614.74
x40 <- 6122.52
x100 <- 34939.95
```


## Read in your data
```{r}
# NOTE: make sure your data file is in the appropriate format!

# you can either select it directly by calling the file name
dat <- read.csv("20140519_CountsExample.csv")

# or by using Windows Explorer
dat <- read.csv(file.choose())

dat <- as.matrix(dat)
```

## Estimating cellular density
```{r}
# Pulls out sample names in file and includes them in a vector
samples <- unique(dat[,2])

# create an empty vector for output
output <- c()

# Calculate cells per mL (cpm) per sample
for (i in 1:length(samples)){
  tmp <- dat[dat[,2] == samples[i],]
	ids <- tmp[1,2:3]
	counts <- tmp[,4:ncol(tmp)]
	fields <- as.numeric(counts[12])
		fields[fields == 20] <- x20
		fields[fields == 40] <- x40
		fields[fields == 100] <- x100
	cpm <- mean(as.numeric(counts[1:10]))/as.numeric(counts[11])*
		(1-as.numeric(counts[13]))*1000*fields
	output <- rbind(output,c(ids,cpm))
}

# View for loop data
output

# Save output on harddrive.
write.csv(output, file = "date_output.csv")
```
